---
title: 'Day 10: Collecting data from the internet using APIs'
output:
  html_document:
    toc: true
    toc_depth: 4    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```
## Do this before class

Solve chapters *Downloading Files and Using API Clients*, *Using httr to interact with APIs directly* and *Handling JSON and XML* på [DataCamp](https://www.datacamp.com/enterprise/mt5013-ht19/).

Read: Chapters 1-3 in [An introduction to APIs](https://zapier.com/learn/apis/)

Resources: 

* [ProgrammableWeb](https://www.programmableweb.com/) contains a catalog of open APIs.
* [Öppna data och PSI](https://oppnadata.se/) contains a catalog of Swedish resources.

## During class

### a-pi

Write a function
```{r, eval = FALSE}
get_pi <- function(start, numberOfDigits){
    ...
}
```
that calls [a-pi](https://pi.delivery/#apipi_get) and returns the digits of $\pi$ from `start` to `start + numberOfDigits - 1`. A sample call is <https://api.pi.delivery/v1/pi?start=1000&numberOfDigits=5>.

### TimeEdit

A TimeEdit-schedule can be obtained in JSON-format by changing `.html` to `.json` in the url, try in a web-browser with 
<https://cloud.timeedit.net/su/web/stud1/ri107455X48Z06Q5Z16g3Y05y5006Y48Q02gQY6Q55727.html>

**Part 1**: We can use `GET` to import it to R

```{r}
library(httr)
schema_response <- GET("https://cloud.timeedit.net/su/web/stud1/ri107455X48Z06Q5Z16g3Y05y5006Y48Q02gQY6Q55727.json")
schema_json <- content(schema_response, "text")
```

The result can be explored with `jsonedit` in the package `listviewer`
```{r}
library(knitr)
library(tidyverse)
library(jsonlite)
library(listviewer)
jsonedit(schema_json)
```


We convert it with  `fromJSON` and choose the  `reservations`
```{r}
schema_df <- fromJSON(schema_json)$reservations
```


The result is now a `data.frame` with six columns, where the last column contains a vector in each cell. In order to extract elements from this column, we use  `mutate` in combination `map_chr`. The family of `map`-functions comes from the `purrr` package which is part of the `tidyverse`, more abot them later. Here `map_chr(columns, 1)` corresponds to `sapply(columns, function (x) x[1])` in base-R.

```{r}
schema_df %>% mutate(sal = map_chr(columns, 3), 
                     kurs = map_chr(columns, 1), 
                     tid = paste(starttime, endtime, sep = " - ")) %>% 
    select(kurs, datum = startdate, tid, sal) %>% 
    kable()
```


Fun fact: The schedule of our github page is based on such a TimeEdit API call - see [`schedule.html`](https://github.com/mt5013-ht19/mt5013-ht19.github.io/blob/master/schedule.Rmd), which makes it easy to adapt the page each year the course is given.


**Part 2**: The schedule for room 14 during the 2nd part of the winter semester can be found on

<https://cloud.timeedit.net/su/web/stud1/ri107455X28Z07Q5Z76g0Y05y5076Y31Q09gQY6Q55777.html>

* What teacher spends most time teaching in room 14 during the 2nd part of the winter semester?


### SCB

If you generate a table at [Statistikdatabasen](http://www.statistikdatabasen.scb.se/), you will find a link "API för denna tabell" that gives an url and a query to be made by POST in order to fetch the table. Try fetching a table with `httr::POST`, the query should be placed in the `body`.

Note: at the end of the query you may change `"format": "px"` to `"format": "json"` in order to get a reply in JSON-format. Fetch a table, examine its structure and try to extract a suitable `data.frame`. See [`SCB_KPI_API.R`](https://github.com/MT5013-HT19/Class_files/blob/master/SCB_KPI_API.Rmd) for a simple example.

Note: A dedicated R package [`pxweb`](https://cran.r-project.org/web/packages/pxweb/pxweb.pdf) exists for querying the SCB database through the API in a slightly more comfortable way. See the [vignette](https://htmlpreview.github.io/?https://github.com/rOpenGov/pxweb/blob/test/vignettes/pxweb.html) for a demonstratio.

### Latest day at Bromma

The last days hourly temperatures (parameter 1) at Bromma (station 97200) can be fetched from SMHI by  (switch `xml` for `json` if you want to change format)

```{r}
temp_response <- GET("https://opendata-download-metobs.smhi.se/api/version/1.0/parameter/1/station/97200/period/latest-day/data.xml")
http_type(temp_response)
```

We extract the XML-content by

```{r}
library(xml2)
temp_xml <- read_xml(temp_response)
class(temp_xml)
```

The structure can be viewed by opening 

<https://opendata-download-metobs.smhi.se/api/version/1.0/parameter/1/station/97200/period/latest-day/data.xml> 

in your web-browser. We see that temperatures can be found in  `XPATH` `"/metObsSampleData/value/value"`:

```{r}
xml_ns_strip(temp_xml) # Överkurs
xml_find_all(temp_xml, "/metObsSampleData/value/value")
```
(use `xml_text` to get the values).


* Also extract the times.
* Try with another station/parameter (see [SMHI](https://opendata-download-metobs.smhi.se/explore/#)). Also try the JSON-format.

### Systembolag

Systembolaget's API uses XML, the list of stores from HW4 can be fetched by
```{r}
stores_response <- GET("https://www.systembolaget.se/api/assortment/stores/xml")
http_type(stores_response)
```

We extract XML with

```{r}
stores_xml <- read_xml(stores_response)
```
and look at the first with

```{r}
xml_find_first(stores_xml, "/ButikerOmbud/ButikOmbud")
```
In order to extract the names we may use
```{r}
xml_find_all(stores_xml, "//Namn")[1:10]
```

evidently, not all stores have names. 

* Convert the result to a `data.frame`.




