---
title: 'Day 2: Tidyverse: Basic `dplyr` and `ggplot2`.'
output:
  html_document:
    toc: true
    toc_depth: 4    
---

## Do this before class

Read [R4DS](http://r4ds.had.co.nz/) chapters 3.1-3.6, 5.1-5.5

Complete assignments *Data wrangling* and *Data visualization* (first two chapters of Introduction to the Tidyverse) at [DataCamp](https://www.datacamp.com/enterprise/mt5013-ht18/assignments).

## Class activities

Start by creating a new R-project "Classroom" that you will use for your class activities. Some activities will require data or scripts from the repo [`Class_files`](https://github.com/MT5013-HT18/Class_files), we therefore recommend that you clone this into a subfolder of your `Classroom` directory. Now create a new R Markdown document `Class1.Rmd` where you will do your work for this class.

# Systembolaget's assortment

[Systembolaget](https://www.systembolaget.se/)'s assortment of beverages from 2018-10-18 is available in the file [`Class_files/systembolaget2018-10-08.csv`](https://github.com/MT5013-HT18/Class_files/blob/master/systembolaget2018-10-08.csv). It is downloaded from [Systembolaget's public API](https://www.systembolaget.se/api) and saved in csv-format by the script [`Class_files/Systembolaget.R`](https://github.com/MT5013-HT18/Class_files/blob/master/Systembolaget.R). Load its contents by

```{r, eval = FALSE}
library(tidyverse)
Sortiment_hela <- read_csv("Class_files/systembolaget2018-10-08.csv")
```


### Exercises (training of `arrange`, `filter`, `mutate`, `select`, `%>%`)

* The variable  `Alkoholhalt` (alcohol by volume) has been classified as `character` by `read_delim`, since it contains a percent sign. Convert it to numeric using `mutate` by first removing the percent sign (e.g. with `gsub`) and then transform with `as.numeric`.

* A few wines have been labelled as `Röda` and `Vita` instead of `Rött vin` (red wine) respektive `Vitt vin` (white wine) in the `Varugrupp` (group of products) column. Fix this, e.g. by using `mutate` and `ifelse`.

* What beverage has the highest `PrisPerLiter`? Display the answer (the `Namn` of the beverage) as dynamically coded in the text body of your  `.Rmd`-document.

* Create a new data frame `Sortiment_ord` with the regular product range (where `SortimentText` equals `Ordinarie sortiment`). Make a table (with `kable` from the `knitr`-library) of the 10 most expensive (`PrisPerLiter`) beverages from this range. Use `select` to select suitable columns for the table.

* if you have not already done so, write the code from the previous excercise using a sequence of pipes (`%>%`).

### Exercises (training of `ggplot`, `geom_point`, `geom_line`, `facet_wrap`)

For the regular product range in `Sortiment_ord`
 
* Plot `PrisPerLiter` against `Alkoholhalt`, color the points by `Varugrupp` and consider using a log-scale for `PrisPerLiter`.
* Plot `PrisPerLiter` (possibly on a log-scale) against `Varugrupp`. Consider `coord_flip` to improve readability.
* For the  groups of products `c("Vitt vin", "Rött vin", "Rosévin", "Mousserande vin")` of vintage (`Argang`) 2010-2017, plot `PrisPerLiter` against `Argang`. Try both using  a `facet` for each group and coloring by group in the same facet.

### Further excercises

* Use your imagination and keep exploring the data.

# Film events

[Stockholm international film festival](https://www.stockholmfilmfestival.se/en) just started. In [`Class_files/Film_events_2018-11-07.csv`](https://github.com/MT5013-HT18/Class_files/blob/master/Film_events_2018-11-07.csv) you will find their event schedule as of this morning.

### Exercises (training of `arrange`, `filter`, `mutate`, `select`, `%>%`)

* What films are already sold out (for all screenings)?
* What venue screens the most number of (unique) films?
* Plot the proportion of sold out events for each day of the festival.

# Olympic winter medals 

The file [`Class_files/Winter_medals2018-09-26.csv`](https://github.com/MT5013-HT18/Class_files/blob/master/Winter_medals2018-09-26.csv) contains the number of medals per country and olympic year at the winter olympics since 1980 together with the total population of the country. The data set is scraped from Wikipedia using the script [`Class_files/Winter_medals.R`](https://github.com/MT5013-HT18/Class_files/blob/master/Medal_table.R) which contains more information, in particular on countries that has been split or joined during the period.

Load the file using
```{r, eval = FALSE}
winter_medals <- read_csv("Class_files/Winter_medals2018-09-26.csv")
```


### Exercises (training of `arrange`, `filter`, `mutate`, `select`, `%>%`)

* Create a variable column `medals_per_mill`, the number of medals per million inhabitants.
* Print a table of the 10 most successful countries, by `medals_per_mill`, during the 2018 Winter olympics.

### Exercises (training of `ggplot`, `geom_point`, `geom_line`, `facet_wrap`)

* Plot the total number of medals against year for Sweden, Norway and Finland in the same figure and separate the countries with a suitable aesthetic (see `?geom_point` for a list of aesthetics `geom_point` understands).
* Plot the number of Swedish gold, silver and bronze medals against year (bonus: color the points in gold/silver/bronze).
* As previous excercise, but with a "`facet`" for each of Sweden, Norway and Finland.

# Gapminder 

Use `ggplot` to recreate (static versions) of some figures from [Hans Rosling's talks](https://www.ted.com/speakers/hans_rosling). Data is available in  package `gapminder`.
