---
title: 'Dag 5: Tidy data'
output:
  html_document:
    toc: true
    toc_depth: 4    
---

```{r setup, echo=FALSE, results="hide", message=FALSE}
library(tidyverse)
```

## Do this before class

Read [R4DS](http://r4ds.had.co.nz/) chapter 12.

Solve *Introduction and exploring raw data* and *Tidying data* of the *Cleaning Data in R* course at [DataCamp](https://www.datacamp.com/enterprise/mt5013-ht19/assignments).

## During class

* The file [`Class_files/Statistikdatabasen_2019-11-14 23_25_40.csv`](https://github.com/MT5013-HT19/Class_files/blob/master/Statistikdatabasen_2019-11-14%2023_25_40.csv) contains records of adults visiting dental care by sex, municipality and year (downloaded from [The National Board of Health and Welfare dental health records](https://sdb.socialstyrelsen.se/if_tandhalsa)). Read it with
```{r, eval = FALSE}
dental_data <- read_csv2("Class_files/Statistikdatabasen_2019-11-14 23_25_40.csv", skip = 1, n_max = 580)
```
(why `skip = 1` and  `n_max = 580`?) and use `gather` to convert it to long ("tidy") format.

* The file [`Class_files/Statistikdatabasen_2018-01-23 15_04_12.csv`](https://github.com/MT5013-HT19/Class_files/blob/master/Statistikdatabasen_2018-01-23 15_04_12.csv) contains records of suicides and death totals by sex, county and year (from [The National Board of Health and Welfare](https://sdb.socialstyrelsen.se/if_dor/val.aspx)). Read it with

```{r, eval = FALSE}
data <- read_csv2("Class_files/Statistikdatabasen_2018-01-23 15_04_12.csv", skip = 1, n_max = 80)
```

and plot the proportion of suicides among death totals, by year, for the whole country.

* The file [`Class_files/Statistikdatabasen_2018-01-23 15_39_06.csv`](https://github.com/MT5013-HT19/Class_files/blob/master/Statistikdatabasen_2018-01-23%2015_39_06.csv) contains monthly records of social assistance (ekonomiskt bistånd). Explore it in a spread sheet and transform to  "tidy format" containing the variable average payment per houshold (`Utbetalt ekonomiskt bistånd tkr` divided by `Antal hushåll`) for each month and county. It may be helpful to join the first two columns using `paste`.

* Examination of the course Matematik I (MM2001) consists a large number of smaller labs/exams with individual codes (Provkoder), see the 
 [course plan](https://sisu.it.su.se/pdf_creator/8940/34903) for a full list. The file [`Class_files/MM2001.csv`](https://github.com/MT5013-HT19/Class_files/blob/master/MM2001.csv) contains results for 100 randomly chosen students exported using and older version of [Ladok](https://sv.wikipedia.org/wiki/LADOK). Each row corresponds to one student (name and id removed). Transform data (or a subset of data) to a tidy format.
 