---
title: 'Day 7: Basic SQL'
output:
  html_document:
    toc: true
    toc_depth: 4    
---

## Do this before class

Complete course *Intro to SQL for Data Science* and chapters *Importing data from databases Part 1 and 2* [DataCamp](https://www.datacamp.com/enterprise/statistical-data-processing-mt5013/assignments). Read the [RSQLite vignette](https://cran.r-project.org/web/packages/RSQLite/vignettes/RSQLite.html).

## During class

There are essentially three ways to send SQL-queries directly from RStudio to a database `mybase`, where we assume `mybase` is an SQLite database stored in file `mybase.sqlite`.

- through `dbGetQuery` after establishing a conection using `dbConnect` as in

```{r, eval = FALSE}
con <- dbConnect(SQLite(), "mybase.sqlite")
dbGetQuery(con, "SELECT * FROM table")
```


- by creating an SQL-chunk in a Markdown document as in

````markdown
`r ''````{r}
con <- dbConnect(SQLite(), "mybase.sqlite")
```

`r ''````{sql, connection = con}
SELECT * FROM table
```
````

- by installing the [RStudio version $\geq$ 1.2](https://blog.rstudio.com/2018/10/02/rstudio-1-2-preview-sql/) which allows the execution of SQL scripts.

We recommend using the first option within an R script file.

### SQL joins

Practise SQL joins at [w3schools](https://www.w3schools.com/sql/sql_join.asp).

### SQL for [Storstockholms Lokaltrafik (SL)](https://sv.wikipedia.org/wiki/Storstockholms_Lokaltrafik) 

The SQLite database [`Class_files/sthlm_metro.sqlite`](https://github.com/mt5013-ht19/Class_files/blob/master/sthlm_metro.sqlite) contains data on the stops of the Stockholm metro at the platform level, which was generated for the 2018 class by [`Class_files/sthlm_metro.R`](https://github.com/mt5013-ht19/Class_files/blob/master/sthlm_metro.R) and originates from [Trafiklab.se](https://www.trafiklab.se/api/sl-hallplatser-och-linjer-2).

- Connect to the database, list the tables and figure out how they relate to eachother.

- Query for the `LineName` from table `Line` where `LineNumber` is 18.
- Query for the number of platforms at each station in descending order.
- Query for the five most southern `StationName`, where position is measured as the average `Latitude` of its `PlatformNumber`.
- Query for the number of stations on `LineNumber` 18.
- Query for all `StationName` on `LineNumber` 18 in alphabetical order.


### Pokemon SQL

The SQLite database `Class_files/pokedex.sqlite` contains the full set of tables from [https://github.com/veekun/pokedex/tree/master/pokedex/data/csv](https://github.com/veekun/pokedex/tree/master/pokedex/data/csv) generated by the file [`Class_files/pokedexDB.R`](https://github.com/mt5013-ht19/Class_files/blob/master/pokedexDB.R).

Close any existing connections with `dbDisconnect` and connect to the database.

- Query for the average `height` and average `weight` of pokemon (`pokemon` table).
- Query for the number of pokemon of less than average `weight`.
- Query for the average weight of `ghost`-type pokemon.
- Query for a table with columns `pokemon_id`, `pokemon_name` and six additional columns giving the pokemons `base_stat`-value in each of categories `hp`, `attack`, `defense`, `special-attack`, `special-defense` and `speed` (see tables `stats` and `pokemon_stats`). You may use SQL to ask for the table in long format and convert it to the final wide format using `spread` in R.


### Connecting to an external database

[Ensembl](http://www.ensembl.org/index.html) hosts a public genomic database with a MySQL server. You can connect to the database holding the human genome using (note the use of `MySQL` rather than `SQLite`)

```{r}
library(RMySQL)
con <- dbConnect(MySQL(), host = "ensembldb.ensembl.org",
                user = "anonymous", password = "", 
                port = 3306)
```
we are now connected to a whole set of databases
```{r}
databases <- dbGetQuery(con, "SHOW DATABASES")
head(databases)
nrow(databases)
```
Yes, plenty of databases. The latest version of the human genome is in `homo_sapiens_core_94_38`, we choose this by
```{r}
dbSendQuery(con, "USE homo_sapiens_core_94_38")
dbListTables(con)
```
Check the [assembly table](https://www.ensembl.org/info/docs/api/core/diagrams/Core.assembly_tables.svg) if you wonder how the tables relate to eachother... We may take a closer look at a table by `DESCRIBE`
```{r}
dbGetQuery(con, "DESCRIBE gene")
```


- What is the primary key in the `gene`-table?
- Query for the total number of protein-coding genes in the human genome (see `biotype` in the `gene`-table).
- Query for the first 10 DNA-sequences in the `dna` table.



